---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# brixtools

The package and site are still under construction...

The goal of brixtools is to make it easier to model non-life insurance data with
deductibles using maximum likelihood methods.

## Installation

You can currently install brixtools from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("kbrix/brixtools")
```

## Example

Included in the package is a simulated data set of 10000 policies. See the documentation for `normal_poisson_policy`. Below is one of the policies with two losses, but only one (reported) claim:

```{r example, tidy=TRUE}
library(brixtools)
occurrence <- lapply(normal_poisson_policy, function(x) unlist(x["occurrence"]))
reported <- lapply(normal_poisson_policy, function(x) unlist(x["reported"]))
set <- intersect(which(occurrence == 2), which(reported == 1))[1]
normal_poisson_policy[[set]]
```

Since the losses (and therefore also claims) within each policy are assumed i.i.d., we can simplify the indexing of the data to ease numerical procedures, see `normal_poisson_x` and `normal_poisson_n`.

The Normal-Poisson log-likelihood is called using `normal_poisson_log_likelihood` and can be used with `brix_simple` to find the optimum yielding the maximum likelihood estimates of the parameters:

```{r}
model <- brix_simple(optimizer = optim,
                     fn = normal_poisson_log_likelihood,
                     X = normal_poisson_x, N = normal_poisson_n,
                     par = rep(1, 3),
                     control = list(fnscale = -1),
                     method = "Nelder-Mead",
                     hessian = TRUE)
summary(model)
```

The true parameter values are 1.5 for the mean and standard deviation for the normal distribution and the mean for the non-thinned Poisson is 0.1. The possiblities for extensions are very large.


